Network Deliberation: The role of network structure in large-scale, internet-enabled, participatory decision-making

### 2022-08-08 UMich/chapters/ch05-discussion.tex
Ch05 revisions.

In the preceeding chapters, I have approached the above question using observational study, numerical simulation, and experiment.
A combination of literature reivew and obersvational analysis of English-language Wikipedia (\ref{chap:wp-prod-perf}) identified a common theme among successful examples: small, tightly-knit groups, interconnected through mutual membership: network deliberation.
Subsequent work investigates simple models of network deliberation to better understand how and why it helps large groups effectively self-govern.
These models are not intended to capture all properties of real-world network deliberation, but rather to isolate and investigate particular factors, such as social influence and network efficiency.
Both numerical simulations (Chapter \ref{chap:abm}) and experiment (Chapter \ref{chap:experiment}) find evidence that network deliberation can improve the effectiveness of large deliberations relative to conventional single-group deliberation.
In particular, these studies both identify a protective effect against the negative consequences of social influence, such as information cascades.
As such, network deliberation may be especially beneficial in settings with a high level of social influence.

\section{Future Work: Deliberative Settings}
While the studies presented here confirm the observation that network deliberation sometimes contributes to the success of large-scale deliberation, there are many factors that can influence the deliberative process, and there remains much work to be done to understand how these factors interact with network deliberation.

@@

\subsection{Facilitation and Moderation}
When all members are equal participants in a deliberation, facilitation and moderation are a challenge.
In the experiment reported in Chapter \ref{chap:experiment}, facilitation was achieved through discussion prompts.
Moderation was performed by an experimenter, but the discussion remained respectful and no moderation actions were required.
An active facilitator might contribute subtantially to the quality of a deliberation, but there are many potential ways to choose a facilitator from among a group of peers.
Similarly, moderation is incredibly important for productive deliberation on sensitive topics, and it is not immediately obvious how moderation might work in a network deliberation environment.

@@

@@

One possible application of network deliberation principles is in the design of organizational structures and procedures.
In fact, small interlocking pods are a feature found in self-managed organizations \cite{laloux_reinventing_2014}.
As the pracitices and structures of such organizations are codified and refined, a better understanding of network deliberation can provide a guide to consructing more effective organizations and institutions.

Similarly, network deliberation emerges naturally in technical settings such as working groups in large collaborations (e.g., WikiProjects), federated social networks (e.g., mastodon), and group chat platforms (e.g., discord).
The deliberation that takes place on these platforms is shaped in part by the protocols and user interaces of the software.
The principles of network deliberation have the potential to inform the design of such protocols and interfaces in a way that minimizes harm and promotes effective interactions.

>>>>

To answer the above question, I have introduced, developed, and applied a model of
large-scale group deliberation: network deliberation.
As a theoretical model, Network deliberation encapsulates common structural pattern
observed across successful large-scale participatory projects:
a network structure built from small interlocking pods.
As with all theoretical models, many parameters are left variable.
In the case of network deliberation, these parameters include:
pod size, pod assignment strategy, and the behavioral/social dynamics of deliberation.
This model can thus be used to analyze how such parameters influence the success of large-scale
participatory projects while providing a common theoretical framework for making comparisons
between different studies.

In the preceeding chapters, I have shown how the network deliberation model can be
used in the analysis of observational studies, numerical simulation, and experiment.
A combination of literature reivew and obersvational analysis of English-language Wikipedia
(\ref{chap:wp-prod-perf}) identified a common theme among successful examples:
small, tightly-knit groups, interconnected through mutual membership: network deliberation.
In Wikipedia for example, articles and their talk pages act as pods,
creating venues for small group interactions, with interlocks created by editors who contribute
to multiple articles.
Subsequent work investigates simple models of network deliberation to better understand
how and why it helps large groups effectively self-govern.
Both numerical simulations (Chapter \ref{chap:abm}) and experiment (Chapter \ref{chap:experiment})
find evidence that network deliberation can improve the effectiveness of large deliberations
relative to conventional single-group deliberation.
In particular, these studies both identify a protective effect against the negative
consequences of social influence, such as information cascades.
As such, these studies of network deliberation suggests an interlocking pod structure may be especially
beneficial in settings susceptible to information cascades, such as those with a high
level of social influence.

\section{Future Work: Deliberative Settings}
While the studies presented here confirm the observation that network deliberation
sometimes contributes to the success of large-scale deliberation,
there are many factors that can influence the deliberative process.
It is difficult to predict whether or how these factors interact with network deliberation
and there remains much work to be done.

@@

\subsection{Facilitation and Moderation}
When all members are equal participants in a deliberation, facilitation and moderation are
a challenge.
In the experiment reported in Chapter \ref{chap:experiment}, facilitation was achieved
through discussion prompts.
Moderation was performed by an experimenter, but the discussion remained respectful and no
moderation actions were required.
An active facilitator might contribute subtantially to the quality of a deliberation,
but there are many potential ways to choose a facilitator from among a group of peers.
Similarly, moderation is incredibly important for productive deliberation on sensitive
topics, and it is not immediately obvious how moderation might work in a network
deliberation environment.
Possibilities might include ad-hoc moderators, or bot/AI-asisted moderation.

@@

\subsection{Next Steps}
The future work needed most immediately is replication of experimental studies of network
deliberation in various contexts.
Such replications can test the robustness and generalizability of the results reported here.
It is also desirable to develop experimental platforms that can accommodate emergent alternatives
that arise during the deliberation, as such alternatives are often preferred to initial proposals
\cite{salganik_wiki_2015}.
The question of the role of structural efficiency also remains.
Our numerical studies found that the effects of varying structural efficiency were much smaller
than those caused by adopting network vs. conventional deliberation,
but we have only evaluated structurally efficient networks empirically.

Large-scale empirical studies present practical challenges, such as recruiting and providing incentives.
One possible approach for overcoming these challenges in future studies is the development
of a publicly-available platform that can be adopted by organizations for practical use,
while simultaneously providing empirical data for study.

@@

One possible application of network deliberation principles is in the design of
organizational structures and procedures.
In fact, small interlocking pods are a feature found in self-managed organizations
\cite{laloux_reinventing_2014}.
As the practices and structures of such organizations are codified and refined,
a better understanding of network deliberation can provide a guide to constructing
more effective organizations and institutions.
Simply recognizing the effectiveness of network deliberation structure suggests dividing
larger teams into pods and ensuring that individuals are active in multiple pods or
periodically rotate through them.
Similarly, our studies suggest there may be benefits to having multiple pods working on the
same tasks in parallel, particularly in the early stages of ideation,
when the exploration/exploitation tradeoff is focused more on exploration.
Furthermore, as the conesequences of parameters such as pod size and assignment methods are
better understood, this understanding may inform the structure and composition of teams.

Similarly, network deliberation emerges naturally in technical settings such as working
groups in large collaborations (e.g., WikiProjects), federated social networks
(e.g., mastodon), and group chat platforms (e.g., discord).
The deliberation that takes place on these platforms is shaped in part by the protocols
and user interfaces of the software.
The principles of network deliberation have the potential to inform the design of such
protocols and interfaces in a way that minimizes harm and promotes effective interactions.
The success of network deliberation favors protocols that allow individuals to participate
in multiple contexts, e.g., forums, microblogging servers, chat rooms.
Similarly, our findings suggest the importance of being able to subdivide larger contexts
into smaller pods.
